Title Goes Here

### 2017-09-14 analysis/paper/saa2018_general_sessions.Rmd
optimum number of topics for SAA2018 general abstracts

```{r eval=FALSE}
# read in the abstracts
general <- read_excel("../data/raw_data/General Abstracts.xlsx")
```

@@

```{r eval=FALSE}
# Create a corpus object from the abstracts
general_corpus <- tm::Corpus(tm::DataframeSource(general), 
                             readerControl = list(reader = tm::readTabular(mapping = m)))

# create a document term matrix
general_tm <- tm::DocumentTermMatrix(general_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    minWordLength = 2, 
                      s                              removeNumbers = TRUE, 
                                                    removePunctuation = TRUE))
```

```{r eval=FALSE}
# Use TF-IDF to wieght terms and remove rare ones
term_tfidf <- 
  tapply(general_tm$v/slam::row_sums(general_tm)[general_tm$i], 
         general_tm$j, mean) * log2(tm::nDocs(general_tm)/slam::col_sums(general_tm > 0))

summary(term_tfidf)
# Median =  0.08989
```

```{r eval=FALSE }
## Keeping the rows with tfidf >= median
general_reduced_dtm <- general_tm[,term_tfidf >= summary(term_tfidf)[3]]
summary(slam::col_sums(general_reduced_dtm))
# save
saveRDS(general_reduced_dtm, "../data/derived_data/general_reduced_dtm.rds")
```

```{r}
general_reduced_dtm <- readRDS("../data/derived_data/general_reduced_dtm.rds")
```

@@

```{r eval=FALSE }
system.time(fitted_many <- 
              lapply(seqk, function(k) topicmodels::LDA(general_reduced_dtm, 
                                                        k = k,
                                                        method = "Gibbs",
                                                        control = list(burnin = burnin,
                                                                       iter = iter, 
                                                                       keep = keep) )))

saveRDS(fitted_many, "../data/derived_data/fitted_many.rds")
```

```{r}
fitted_many <- readRDS("../data/derived_data/fitted_many.rds")
```

```{r}
# extract logliks from each topic
logLiks_many <- lapply(fitted_many, function(L)  L@logLiks[-c(1:(burnin/keep))])

# compute harmonic means
hm_many <- sapply(logLiks_many, function(h) harmonicMean(h))

# to know the optimal number of topics:
optimal_number_of_topics <- seqk[which.max(hm_many)]

# plot it
library(ggplot2)
lda_plot <- ggplot(data.frame(seqk, 
                             hm_many), 
                  aes(x=seqk, 
                      y=hm_many)) + 
  geom_path(lwd=1.5)  +
  xlab('Number of Topics') +
  ylab('Harmonic Mean') +
  geom_vline(xintercept = seqk[which.max(hm_many)],
             colour = "red") +
  annotate("text",
           x = 55, 
           y = -150000, 
           label = paste("The optimal number of topics is", 
                         seqk[which.max(hm_many)])) +
  theme_bw() +
  ggtitle(expression(atop("Latent Dirichlet Allocation Analysis of SAA General Abstracts", atop(italic("How many distinct topics in the abstracts?"), ""))))
```

>>>>

```{r eval=FALSE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")
```

@@

```{r eval=FALSE}
# Create a corpus object from the abstracts
general_2018_corpus <- tm::Corpus(tm::DataframeSource(general_2018), 
                             readerControl = list(reader = tm::readTabular(mapping = m)))

# create a document term matrix
general_2018_tm <- tm::DocumentTermMatrix(general_2018_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    minWordLength = 2, 
                                                    removeNumbers = TRUE, 
                                                    removePunctuation = TRUE))

```

```{r eval=FALSE}
# Use TF-IDF to wieght terms and remove rare ones
term_2018_tfidf <- 
  tapply(general_2018_tm$v/slam::row_sums(general_2018_tm)[general_2018_tm$i], 
         general_2018_tm$j, mean) * log2(tm::nDocs(general_2018_tm)/slam::col_sums(general_2018_tm > 0))

summary(term_2018_tfidf)
# Median =   0.08769
```

```{r eval=FALSE }
## Keeping the rows with tfidf >= median
general_2018_reduced_dtm <- general_2018_tm[,term_2018_tfidf >= summary(term_2018_tfidf)[3]]
summary(slam::col_sums(general_2018_reduced_dtm))
# save
saveRDS(general_2018_reduced_dtm, "../data/derived_data/general_2018_reduced_dtm.rds")
```

```{r}
general_2018_reduced_dtm <- readRDS("../data/derived_data/general_2018_reduced_dtm.rds")
number_of_abstracts <- general_2018_reduced_dtm$nrow
```

@@

```{r eval=FALSE }
system.time(fitted_many_2018 <- 
              lapply(seqk, function(k) topicmodels::LDA(general_2018_reduced_dtm, 
                                                        k = k,
                                                        method = "Gibbs",
                                                        control = list(burnin = burnin,
                                                                       iter = iter, 
                                                                       keep = keep) )))

saveRDS(fitted_many_2018, "../data/derived_data/fitted_many_2018.rds")
```

```{r}
fitted_many_2018 <- readRDS("../data/derived_data/fitted_many_2018.rds")

# how many abstracts, if we just start from here?
how_many_abstracts <- fitted_many_2018[[1]]@Dim[1]
```

```{r}
# extract logliks from each topic
logLiks_many_2018 <- lapply(fitted_many_2018, function(L)  L@logLiks[-c(1:(burnin/keep))])

# compute harmonic means
hm_many_2018 <- sapply(logLiks_many_2018, function(h) harmonicMean(h))

# to know the optimal number of topics:
optimal_number_of_topics_2018 <- seqk[which.max(hm_many_2018)]

# plot it
library(ggplot2)
lda_plot <- ggplot(data.frame(seqk, 
                             hm_many_2018), 
                  aes(x=seqk, 
                      y=hm_many_2018)) + 
  geom_path(lwd=1.5)  +
  xlab('Number of Topics') +
  ylab('Harmonic Mean') +
  geom_vline(xintercept = seqk[which.max(hm_many_2018)],
             colour = "red") +
  annotate("text",
           x = 55, 
           y = -150000, 
           label = paste("The optimal number of topics is", 
                         seqk[which.max(hm_many_2018)])) +
  theme_bw() +
  ggtitle("Latent Dirichlet Allocation Analysis of SAA General Abstracts",
          subtitle = paste0("How many distinct topics are present in the general abstracts (n = ", how_many_abstracts, ") for 2018?"))
```
